### 2016-10-04 article/ms.tex
Init article, plot correlations between nodes

%% Copyright 2016 Andrew R. Casey.  All rights reserved.

\documentclass[preprint]{aastex}

\usepackage{amsmath}
\usepackage{bm}

\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}\newcommand{\giturl}{UNKNOWN}}

\newcommand{\acronym}[1]{{\small{#1}}}

\newcommand{\project}[1]{\textsl{#1}}

\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\logg}{\log g}
\newcommand{\feh}{[\mathrm{Fe/H}]}

\begin{document}
\title{The Gaia-ESO Survey: Combined estimates from multiple stellar analysis pipelines}

\author{
    Andrew~R.~Casey\altaffilmark{1}
}

\begin{abstract}
  Some stuff
\end{abstract}

\keywords{stars: fundamental parameters --- stars: abundances
% Clear rest of page
\clearpage}

\section{Introduction} 
\label{sec:introduction}

\section{Data}
\label{sec:data}

% Description of the setups

% Summary of the ges type and SNR of the data.

% Description of target sources included: MW, Clusters, CoRot, K2, etc.

% Blind test.

\section{Methods}
\label{sec:methods}

% Multiple analysis strategy: why

% primary advantage and disadvantage to multiple analysis strategy

% Description of two parameter rounds: stellar parameters, abundances

% What's reported by each node? Flags? Consistency of flags?

% Description of pipelines in appendix

% Commonalities between nodes: model atmospheres, line list, normalized spectra, initial params, spectrum library, xi calibration if necessary.

% EW method and comparisons -- into an appendix?.

% What do the nodes provide? Most parameters, flags, etc.

% CONTINUE from here:
% Stellar parameter determination

% Stellar parameter homogenisation

% Abundance determination

% Abundance homogenisation

\section{Discussion}
\label{sec:discussion}

% Is the multiple-analysis strategy justified?
% --> Take any one node, infer their uncertainties as a function of SNR using point-wise correlations (e.g., no benchmarks).
% --> What goes wrong for each node?

% Precision, accuracy, systematics.
% EW

% Analysis choices: we give same line list but not all are used.
%                   or how they are used.

\section{Conclusions}
\label{sec:conclusions}

% This should be a "lessons learned" list.

% The final paragraph should contain a succint forward outlook of how good or bad an idea this was.

\end{document}